---
title: "Reproducible Research Course Project 1"
author: "Teodor Gottfried"
date: "Januar 2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(xtable)
Sys.setlocale("LC_TIME", "English") # for non English locale 
activity <- read.csv(choose.files())
library(lubridate)
activity$date<-ymd(activity$date) # make date column as date
```
    
## Number of steps per day

```{r sum n steps}
StepsPerDay <- activity %>% 
    group_by(date) %>% 
    summarise(total_sum = sum(steps))
```
     
### Histogram of the total number of steps taken each day


```{r histogram steps}
hist_plot <- ggplot(StepsPerDay, aes(total_sum))
hist_plot+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", fill="green")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,18,2))
```
     
###Median and mean on original set
```{r median and mean}
ms <- StepsPerDay %>% 
    summarise(`median steps:`=median(total_sum, na.rm = TRUE),  `mean steps:`=mean(total_sum, na.rm = TRUE))
knitr::kable(ms, caption = "Median and mean of steps", format = "html")

```
     
##What is the average daily activity pattern?
   
###Time series plot of the 5 minute interval (x) and averaged number of steps taken averaged across all days (y)
```{r plot steps}
StepsPerTime <- activity %>% 
    group_by(interval) %>% 
    summarise(steps = mean(steps, na.rm=TRUE))
StepsPerTime$time <- StepsPerTime$interval/100
steps_plot <- ggplot(StepsPerTime, aes(time, steps))
steps_plot+geom_line(col="darkgreen")+ggtitle("Average steps per time interval")+xlab("Time")+ylab("Steps")
```
     
###5-minute interval (on average across all the days) with the maximum number of steps
```{r five min inter}
five <- StepsPerTime %>% select(time, steps) %>% filter(steps==max(StepsPerTime$steps))
knitr::kable(five, caption = "5-minute interval", format = "html")
```
##Imputing missing values
###Total number of missing values
```{r total n na}
activity %>% filter(is.na(steps)) %>% 
    summarize(`missing values:` = n())
```
     
###Replace missing values 
complete missing values by rounded values of the average 5-minute interval
```{r mis value}
activity$CompleteSteps <- ifelse(is.na(activity$steps), round(StepsPerTime$steps[match(activity$interval, StepsPerTime$interval)],0), activity$steps)
```
      
###New dataset that is equal to the original dataset but with the missing data filled in
```{r full set}
activityFull <- data.frame(steps=activity$CompleteSteps, interval=activity$interval, date=activity$date)
head(activityFull)

```
###Histogram of the total number of steps taken each day with missing data filled in
```{r data full}
StepsPerDayFull <- activityFull %>% 
    group_by(date) %>% 
    summarise(total_sum = sum(steps))
```
     
```{r histogram full}
hist_plot_full <- ggplot(StepsPerDay, aes(total_sum))
hist_plot_full+geom_histogram(boundary=0, binwidth=2500, col="darkblue", fill="blue")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,18,2))
```
   
###Median and mean on full set
```{r}
msf <- StepsPerDayFull %>% 
    summarise(`median steps:`=median(total_sum, na.rm = TRUE),  `mean steps:`=mean(total_sum, na.rm = TRUE))
knitr::kable(ms, caption = "Median and mean of steps full", format = "html")
```

##Are there differences in activity patterns between weekdays and weekends?
### Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r new factor}
activityFull$weekday <- weekdays(activityFull$date)
activityFull$DayType <- ifelse(activityFull$weekday=='Saturday' | activityFull$weekday=='Sunday', 'weekend','weekday')
head(activityFull)
```
     
### Two time series plot of the 5-minute interval (x) and the average number of steps taken averaged across weekday days or weekend days (y).
```{r plot steps full}

StepsPerTimeWT <- activityFull %>% 
    group_by(interval,DayType) %>% 
    summarise(steps = mean(steps, na.rm=TRUE))

StepsPerTimeWT$time <- StepsPerTime$interval/100

steps_plot_full <- ggplot(StepsPerTimeWT, aes(time, steps))
steps_plot_full+geom_line(col="red")+ggtitle("Average steps per time interval: weekdays vs. weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(size=12))+facet_grid(DayType ~ .)
```

